Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-05-09 code/rmarkdown/article.Rmd
+ aritcle + figure fixes + use glmms for scenario analyses
Former-commit-id: f0708b88545153350a42f98c90def8abd858deee

@@

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56). The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($geq 80$ %). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`$; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($geq 95$ %). Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures SX and SX; such as _Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three different scenarios (Figure 3). The prioritizations generated under the single-species scenario showed relatively homogeneous selection frequencies across the study area--regardless of the combination of targets used to generate them (Figures 3a, 3b, 3c). On the other hand, the prioritizations generated under the multi-species scenario showed different selection frequencies depending on the targets used to generate them. The prioritization generated using only amount-based targets contained fewer planning units (Figure 3d; $n = `r sum(selections(multi.spp.prioritisations[[1]]))`$) than the prioritizations generated using additional of surrogate-based targets (Figure 3e; $n = `r sum(selections(multi.spp.prioritisations[[2]]))`$) or genetic-based targets (Figure 3f; $n = `r sum(selections(multi.spp.prioritisations[[3]]))`$). The prioritizations generated under the multi-species with cost scenario had different spatial distributions to the previous two scenarios. These prioritizations tended to concentrate conservation effort towards areas with higher elevation. Similar to the second scenario, the amount-based prioritization contained less panning units (Figure 3g; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[1]]))`$) than either the surrogate-based (Figure 3h; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[2]]))`$) or the genetic-based prioritizations (Figure 3i; $n = `r sum(selections(multi.spp.prioritisations.with.cost[[3]]))`$).

Generally, the prioritizations generated using amount-based targets met the targets for adaptive genetic variation for most species associated with adaptive loci (Figure 3). Among the three scenarios, the amount-based prioritization in the single-species scenario had the lowest proportion of species with targets met (`r ` % of species with adaptive loci; Figure 3a). In the other two scenarios--the multi-species (`r ` % species preserved with adaptive loci; Figure 3c) and multi-species with cost (`r ` % species preserved with adaptive loci; Figure 3e) scenarios--the amount-based prioritizations successfully preserved the adaptive variation for most of the species. The addition of surrogate- and genetic-based targets did little to improve the proportion of species with their targets met for adaptive genetic variation (Figures 3c and 3e). These results suggest that the main gradients of intra-specific adaptive variation can be sufficiently preserved using suitable amount-based targets.

The prioritizations generated using additional surrogate-based targets met the neutral genetic variation targets for more species than the prioritizations generated using just amount-based targets (Figure 3). The biggest difference between the prioritizations generated using additional surrogate-based targets and just the amount-based targets was in the single-species scenarios (`r ` % and `r ` % of species meeting genetic variation targets respectively). In the other two scenarios, the amount-based prioritizations tended to perform slightly better (`r ` % of species with targets met), but were outperformed by prioritizations with additional surrogate-based targets. Although the surrogate-based targets tended to fulfill the neutral genetic variation targets for most species, they captured only a very small proportion of neutral genetic variation for two species in particular (_Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest that surrogate-based targets can substantially improve the proportion of neutral genetic variation secured in a prioritization.

## Discussion
<!-- introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this hypothesis using moderately realistic planning scenarios. Overall, we found that environmental and geographic variables are effective surrogates for a few species when developing prioritizations. Furthermore, when developing prioritizations under various planning scenarios, we found that the addition of environmental and geographic targets resulted in prioritizations that sufficiently represented the genetic variation of a greater number of species. Our results suggest that while the environmental and geographic variables have little benefit for most species, they can make a large difference for some species.

<!-- consequences for conservation planning -->
+ Typically, conservation planning exercises do not consider intra-specific variation when designing prioritizations. 
+ our results show that for most species, a sufficient proportion of their intra-specific variation can be secured using suitable amount-based targets alone.
+ in particular, we found that using suitable amount-based targets in a multi-species context with comprehensive set of target species resulted in a prioritization that adequately represented the adaptive genetic variation for nearly all the target species (panel c; Figure 3).
+ however, for some species, such a prioritization secures only a small proportion of their genetic variation.
+ we found that for some species, prioritizations generated using amount-based targets only secured a small amount of their neutral genetic variation (panels b, d, e; Figure 3).
+ in planning exercises where the decision maker does not have access to genetic data on their target species across the extent of the study area--ie. nearly all real world planning scenarios--the decision maker will not know which species are poorly represented in terms of their inta-specific variation.
+ the poorly represented species may be those of particular significant or they may not.
+ the decision maker has to weigh the increased costs of a prioritization that includes environmental and geographic surrogates against the risk that a prioritization that does not include these surrogates fails to represent species of conservation significance.
+ here we found that prioritizations generated using environmental and geographic surrogates have only minor increased costs (compare panels c and d; panels e and f; Figure 3).
+ overall, these results suggest that although surrogates are unnecessary for most species, they can be highly beneficial for a few species, the increased costs associated with using these surrogates is small, so generally using surrogates is a good idea.

<!-- comparison to previous studies that have investigated genetic studies -->
+ Previous studies on this data set have found correlations between environmental and genetic variation.
+ Yet, for most species, prioritizations generated using environmental surrogates did not secure substantially more genetic variation than randomly generated prioritizations with the same number of planning units.
+ Here we discuss some potential reasons for this disagreement.
+

+ This apparent disagreement can be attributed to multiple reasons.

+ These results highlight the importance of using prioritizations to assess the effectiveness of potential surrogates for conservation and not simply measuring the strength of correlation between potential surrogates and genetic variation.

<!-- limitations/future directions -->
We wish to alert our readers to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have  artificially introduced additional spatial auto-correlation into the dataset. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.

### Concluding remarks
The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning. We found that environmental and geographic surrogates are effective. However, they seem to be largely unnecessary in conservation planning scenarios because a comprehensive set of species can be used to achieve a similar result. Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principals of conservation planning.

>>>>

We investigated the benefits of using surrogate-based targets in prioritizations using generalized linear mixed-effects models fit with logit liknks [GLMMs; using the \texttt{lme4 R} package; @469]. We fit a full model to the proportion of genetic variation secured in a given prioritization. This model contained categorical predictor variables to indicate the combination of targets used to generate the prioritizations (amount-only, amount and surrogate targets, and amount and genetic targets), the scenario that was used to generate the prioritization (single-species, multi-species, and multi-species with cost), the type of genetic variation measured (adaptive and neutral), and all interactions between these variables. We subject the full model to a backwards stepwise term deletion routine to assess term significance. We conducted a post-hoc analysis on the minimal adquate model using Tukey contrasts with a Bonferroni correction [using the \texttt{multcomp R} package; @r470].

@@

<!-- surrogacy analysis -->
The environmental and geographic variables were moderately effective surrogates for genetic variation in most species (Figure 2). The relationship between the proportion of adaptive genetic variation secured in a prioritization and the proportion of environment variation it also secured was found to depend on the species for which the prioritization was generated (${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). The slope for each species were all significantly different to zero (minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`; maximum $P `r  pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). The explanatory power of the environmental surrogates were also generally quite high for most species (mean `r round(mean(env.surrogate.spp.DF$r2CU),2)` $R^2$ $\pm$ `r round(sd(env.surrogate.spp.DF$r2CU),2)` SD; Figure S56). The nature of these relationships suggest that most of the adaptive variation for a given species can be secured in a prioritization using moderately high environmental-based targets ($geq 80$ %). Similarly, the proportion of neutral genetic variation in a prioritization was affected by the interaction between species and the proportion of neutral variation secured in the prioritization (${\chi^2}_{`r geo.surrogacy.AOV$df[[4]]`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ = `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). The slope for each species were also significantly different to zero (minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`$; maximum $P = `r pvalString(max(env.surrogate.spp.DF$p.value))`; Table S3). Generally, the explanatory power of geographic variables were higher than that of environmental variables (mean `r round(mean(geo.surrogate.spp.DF$r2CU),2)` $R^2 \pm$ `r round(sd(geo.surrogate.spp.DF$r2CU),2)` SD; Figure S57). Compared to the environmental variables, the slopes of the relationships suggest that very high geographic-based targets are required to generate prioritizations that secure a sufficient proportion of neutral genetic variation for most species ($geq 95$ %). Additionally, all species were associated with prioritizations that secured a large proportion of environmental or geographic variation (> 70 %), yet preserved only a very small proportion of the genetic variation (< 10 %; Figures S55 and S56; such as _Gypsophila repens_ and _Luzula alpinopilosa_). These results suggest while prioritizations generated using surrogate-based targets tend to secure more genetic variation, prioritizations generated using just these targets can result in poor solutions.

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according according to three scenarios using different combinations of targets (Figure 3). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4). The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.4.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`). Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate ($Z_1 = `r scenario.GLHT.2.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables ($Z_1 = `r scenario.GLHT.2.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.2.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.3.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`). We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r scenario.ANOVA.2.DF$statistic[[2]]`; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`). Specifically, both the multi-species ($Z_1 = `r scenario.GLHT.1.DF$statistic[[1]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost ($Z_1 = `r scenario.GLHT.1.DF$statistic[[2]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario. Neither the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r scenario.GLHT.1.DF$statistic[[3]]`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$). These results suggest that surrogate-based targets can substantially improve the proportion of genetic variation secured in a prioritization.

## Discussion
* introduction

* consequences for conservation planning

* outlier detection methods - how does our findings agree with other papers?

* mapping genetic variation - compare with GDMS

* limitations/future directions
We wish to alert our readers to several limitations associated with our analysis. First, the size of the planning units we used (23 $\times$ 25 km) is much larger than typically used in conservation planning scenarios (between 1--10 km^2). We used this resolution because the genetic data was collected at this scale. Whilst we could have interpolated the genetic data to a finer resolution and used smaller planning units, this would have biased our analysis because this process would have  artificially introduced additional spatial auto-correlation into the dataset. Second, we used geographic distances as surrogates for connectivity between planning units. Although connectivity between areas can be modeled using isolation by resistance measures [@r473] derived from landscape data [eg. @r474], they are unlikely to be of use in broad-scale conservation exercises since they need to be tuned using species-specific parameters. Third, we used AFLP data to characterize the main gradients of genetic variation. Whilst next-generation sequencing provides higher resolution genetic information than AFLP data [reviewed in @r480], our methodology would still have utilized only the main gradients of the genetic variation in order to generate prioritizations in a feasible amount of time. Overall, we are confident that the limitations associated with our study are relatively minor.

* Concluding remarks
The overarching aim of our study was to investigate the effectiveness of surrogates in conservation planning. We found that environmental and geographic surrogates are effective. However, they seem to be largely unnecessary in conservation planning scenarios because a comprehensive set of species can be used to achieve a similar result. Conservation planners are therefore urged to use a comprehensive set of species in prioritizations--one of the fundamental principals of conservation planning.

### 2016-05-09 code/rmarkdown/figures.Rmd
+ aritcle + figure fixes + use glmms for scenario analyses
Former-commit-id: f0708b88545153350a42f98c90def8abd858deee

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	left_join(
		data.frame(
			letter=unname(scenario.G.cld$Letters),
			Combination=as.character(names(scenario.G.cld$Letters))
		) %>% mutate(Combination=as.character(Combination)),
		by='Combination'
	) %>% data.frame()

# make plot
suppressWarnings({
	ggplot(aes(x=Prioritisation,y=genetic.held),
		data=scenario.DF) +
		geom_boxplot(position=position_dodge(0.9)) +
		geom_text(aes(x=Prioritisation, y=letter.height, label=letter),
			data=scenario.PDF, position=position_dodge(0.9)) +
		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
			data=scenario.PDF) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits=c(NA, 1.2)) +
		theme_classic() +
		theme(strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white'),
			panel.border = element_rect(color='black', fill=NA),
			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
		geom_abline(intercept=rapr.params.LST[[MODE]]$scenario.analysis$genetic.target,
			slope=0, color='red', linetype='dashed') +
		facet_grid(Context ~ Metric)
})
```

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=5, fig.height=5.5, fig.cap='The effectiveness of prioritizations at securing genetic variation. The top row of panels (a--b) show the performance of single-species prioritizations generated assuming equal costs among planning costs. The middle row of panels (c--d) show the performance of multi-species prioritizations generated assuming equal costs among planning units. The bottom row of panels (e--f) show the performance of multi-species prioritizations generated using opportunity costs. The left column of panels (a, c, d) and the right column of panels (b, d, f) show the proportion of adaptive and neutral genetic variation secured (respectively). Box plots show the effectiveness of prioritizations generated using amount-based targets, amount-based and surrogate-based targets, and amount-based and genetic-based targets for each species. Letters denote significant differences ($P < 0.05$).'}
# prepare data for plotting
scenario.PDF <- scenario.DF %>% group_by(Prioritisation, Context, Metric) %>% 
	mutate(
		letter.height=max(genetic.held, na.rm=TRUE)+0.1,
		Combination=paste0(Prioritisation,'.',Metric,'.',Context)) %>%
	filter(c(TRUE, rep(FALSE, length(Context)-1))) %>% 
	ungroup() %>%
	group_by(Context, Metric) %>% 
	mutate(
		panel.letter=paste0(
			'(',letters[(as.integer(Context)*2)-1 + (as.integer(Metric)-1)],')'
		),
		panel.letter=c(as.character(panel.letter[1]), rep(NA_character_, length(Context)-1))
	) %>% 
	ungroup() %>% 
	data.frame()

# make plot
suppressWarnings({
	ggplot(aes(x=Prioritisation,y=genetic.held),
		data=scenario.DF) +
		geom_boxplot(position=position_dodge(0.9)) +
		geom_text(aes(x=0.6, y=1.15, label=panel.letter),
			data=scenario.PDF) +
		scale_fill_manual(name='Prioritisation',
			values=c('grey80','grey50','grey20')) +
		ylab('Genetic variation secured (%)') +
		xlab('') +
		scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1), limits=c(NA, 1.2)) +
		theme_classic() +
		theme(strip.background = element_rect(fill='grey20'),
			strip.text = element_text(color='white'),
			panel.border = element_rect(color='black', fill=NA),
			axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
		facet_grid(Context ~ Metric)
})
```
